Sparse Function-space Representation \\ of Neural Networks

### 2023-05-15 paper/main.tex
mINor

Our approach contrasts to \citet{immer2021improving}, which uses a subset of the full model to try to reduce complexity, a less principled approach to build a sparse model given it ignores the information in the dataset. It is important to highlight that the sparsifying process is independent of the approximate inference technique. The \cref{eq:dual_param} is a result of using \cref{eq-empirical-risk} and a Laplace approximation. More complicated inference techniques, such as variational inference, could be used, but given its simplicity, we only need a trained neural network; we use a Laplace approximation in our experiments. Furthermore, our dual variable view of approximate inference means we do not need to resort to retraining a separate sparse GP model.

@@

\subsection{\our for Continual Learning}
% Definition of CL -> a sequence of tasks 
In the Continual learning setting, the training is divided into $T$ tasks, each with its training data set $\dataset_t = \{(\mathbf{x}_{i}, \mathbf{y}_{i})\}_{i=1}^{n_t}$, which after the task is discarded and cannot be accessed in the later stages of the training. However, rehearsal and function regularization-based models keep a subset of the training data for each task to help the model alleviate forgetting. After each task, we can use our \our to build a compact representation of the neural network and then use that as an additional term to the training loss to regularize the model. Recent success has shown GP-based regularizers perform better than weight space equivalents getting state-of-the-art on typical CL benchmarks \cite{ pan2020continual, rudner2022continual}. The limitation of such approaches is that they must use subset approximations of the posterior, which in practice means randomly selecting some data computing its posterior at only those points and then using that as the regularizer. A more principled approach would be to sparsify the data into a low-rank approximation to ensure information wasn't lost, essentially what the sparse dual variables do.   \todo{this term is replacing the missing data}

An alternative view of the dual variables is that they parameterize approximate likelihoods, as shown in \citep{adam2021dual, khan2017conjugate}. We can rewrite the GP posterior for $q(\vu)$ in terms of the following approximate normal distributions:
\begin{equation}
 q(\vu) \propto \N(0, \MKzz) \prod_{i=1}^n e^{-\frac{\tilde{\beta}_i}{2}(\tilde{y}_i - \va_i^\top \vu)^2 } = \N(0, \MKzz) \, e^{(\bar{\vy} - \vu)^\top \bar{\MSigma}^{-1}(\bar{\vy} - \vu)}  
\end{equation}
The first normals are the approximate likelihoods where $\va_i^\top = \vkzi^\top \MKzz^{-1}$ and $\tilde{y}_i = \hat{\alpha}_i + f_{\vw^*}(\vx_i)\hat{\beta}$,  replacing $\va_i^\top \vu$ with $\vx_i^{\top}\vw$ and we can see the similarity between the sparse GP and a linear model where the kernel matrix determines the prior. To arrive at the MVN Gaussian, we need to rearrange the quadratic terms to match the prior sufficient statistics we present a derivation in [appendix]. Performing conjugate summation of prior and likelihood, we would arrive at the sparse GP posterior at the inducing points. The correlated covariance structure $\MSigma$ and the corresponding $\tilde{\vy}$ are simply a different form of the sparse dual variables,
\begin{equation}
\quad \bar{\MSigma} =  \MKzz \vbeta_\vu^{-1} \MKzz \quad \bar{\vy} = \MKzz \vbeta_{\vu}\valpha_{\vu}.
\end{equation}
Written as above, we can see that the spare dual variables can be interpreted as a sparse MVN representation. We now show how we use this for a regularizer in continual learning with more details in [appendix].

@@

@@

\subsection{Reinforcement learning}
A key challenge in RL is balancing the trade-off between exploration and exploitation.
That is, should the agent select an action that it knows will lead to a high reward (exploitation), or should it
hope to discover new actions which lead to high reward (exploration) by selecting actions that it has not selected before.
In this section, we demonstrate  that our method's principled uncertainty estimates can be used to balance this trade-off in a model-based RL algorithim.

We consider environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expectation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
\end{align}
where $\gamma \in [0, 1)$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics is learned \(f_{\mathbf{w}} \approx \transitionFn\) and then used by a planning algorithm.
A simple approach is to use the learned dynamic model $f_{\mathbf{w}^{*}}$ and maximise the objective in \cref{eq-model-free-objective}.
% \begin{align} \label{eq-greedy}
%   \policy^{\text{Greedy}} &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi).
% \end{align}
However, we can leverage the method in \cref{sec:methods} to obtain a function-space posterior over the learned dynamics $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$,
where $\mathcal{D}$ represents the state transition data set \(\mathcal{D} = \{(s_{i},a_{i}), s_{i+1}\}_{i=0}^{N}\).
Importantly, the uncertainty represented by this posterior distribution can be used to balance the exploration-exploitation trade-off,
using approaches such as posterior sampling \cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013},
\begin{align} \label{eq-posterior-sampling}
  \policy^{\text{PS}} &= \arg \max_{\pi \in \Pi} J(\hat{f}, \pi)
\quad \text{s.t. } \tilde{\transitionFn} \sim q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset),
\end{align}
where a function $\tilde{\transitionFn}$ is sampled from the (approximate) posterior $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$ and used to find a policy as
in \cref{eq-greedy}.
Intuitively, this strategy will explore where the model has high uncertainty, which in turn will reduce the model's uncertainty as data is collected and used to
train the model.
See \cref{sec-rl-appendix} for more details of how we use

\todo{could try to implement Pathwise conditioning in function-space}

See \cref{app:rl} for more details of our experiments.

@@

\nipstitle{
    {\Large Supplementary Material:} \\
    PAPER TITLE GOES HERE}
\pagestyle{empty}

>>>>

Our approach contrasts to \citet{immer2021improving}, which uses a subset of the full model to try to reduce complexity, a less principled approach to build a sparse model given it ignores the information in the dataset. It is important to highlight that the sparsifying process is independent of the approximate inference technique. The \cref{eq:dual_param} are from the neural network definition \cref{eq-empirical-risk}. More complicated inference techniques, such as variational inference, could be used, but given its simplicity, we only need a trained neural network; we use a Laplace approximation in our experiments. Furthermore, our dual variable view of approximate inference means we do not need to resort to retraining a separate sparse GP model.

@@

\subsection{\our for Continual Learning}
% Definition of CL -> a sequence of tasks 
In the Continual learning setting, the training is divided into $T$ tasks, each with its training data set $\dataset_t = \{(\mathbf{x}_{i}, \mathbf{y}_{i})\}_{i=1}^{n_t}$, which after the task is discarded and cannot be accessed in the later stages of the training. However, rehearsal and function regularization-based models keep a subset of the training data for each task to help the model alleviate forgetting. After each task, we can use our \our to build a compact representation of the neural network and then use that as an additional term to the training loss to regularize the model. \todo{this term is replacing the missing data}

An alternative view of the dual variables is that they parameterize approximate likelihoods as was shown in \citep{adam2021dual} we can rewrite the GP posterior in terms of the following approximate normal distributions:
\begin{equation}
 q(\vu) \propto \N(0, \MKzz) \prod_{i=1}^n e^{-\frac{\hat{\beta}_i}{2}(\tilde{y}_i - \va_i^\top \vu)^2 } = \N(0, \MKzz) \, e^{(\tilde{\vy} - \vu)^\top\tilde{\MSigma}^{-1}(\tilde{\vy} - \vu)}  
\end{equation}
The first normals are the marginal likelihoods were $\va_i^\top = \vkzi^\top \MKzz^{-1}$ and $\tilde{y}_i = \hat{\alpha}_i + f_{\vw^*}(\vx_i)\hat{\beta}$,  replacing $\va_i^\top \vu$  wtih $\vx_i^{\top}\vw$ and we can see the similarity between the sparse GP and a linear model where the prior is determined by the kernel matrix. To arrive at the MVN Gaussian we need to rearrange the quadratic terms to match the prior sufficient statistics we present a derivation in [appendix]. Performing conjugate summation of prior and likelihood we would arrive at the sparse GP posterior at the inducing points $\MZ$. The correlated covariance structure $\MSigma$ and the corresponding $\tilde{\vy}$ are simply a different form of the sparse dual variables

@@

\subsection{Reinforcement learning}
A key challenge in RL is to balance the trade-off between exploration and exploitation.
That is, should the agent select an action that it knows will lead to a high reward (exploitation), or should it
hope to discover new actions which lead to high reward (exploration) by selecting actions that it has not selected before.
In this section, we detail how our method's principled uncertainty estimates can be used to balance this trade-off in a model-based RL algorithim.

We consider stochastic environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expectation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
\end{align}
where $\gamma \in [0, 1)$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics is learned \(f_{\mathbf{w}} \approx \transitionFn\) and then used by a planning algorithm.
A simple approach is to use the learned dynamic model $f_{\mathbf{w}^{*}}$ and maximise the objective in \cref{eq-model-free-objective},
\begin{align} \label{eq-greedy}
  \policy^{\text{Greedy}} &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi).
\end{align}
However, we can leverage the method in \cref{sec:methods} to obtain a function-space posterior over the learned dynamics $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$,
where $\mathcal{D}$ represents the state transition data set \(\mathcal{D} = \{(s_{i},a_{i}), s_{i+1}\}_{i=0}^{N}\).
Importantly, the uncertainty represented by this posterior distribution can be used to balance the exploration-exploitation trade-off,
using approaches such as posterior sampling \cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013},
\begin{align} \label{eq-posterior-sampling}
  \policy^{\text{PS}} &= \arg \max_{\pi \in \Pi} J(\hat{f}, \pi)
\quad \text{s.t. } \tilde{\transitionFn} \sim q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset),
\end{align}
where a function $\tilde{\transitionFn}$ is sampled from the (approximate) posterior $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$ and used to find a policy as
in \cref{eq-greedy}.
Intuitively, this strategy will explore where the model has high uncertainty, which in turn will reduce the model's uncertainty as data is collected and used to
train the model.
See \cref{sec-rl-appendix} for more details of how we use

\textbf{Pathwise conditioning in function-space??}
\todo{could try to implement this}

@@

\subsection{Reinforcement Learning}

In section \label{sec-rl} we detailed how our method can be used to

See \label{sec-rl-experiments-appendix} for more details of our experiments.

@@

\nipstitle{
    {\Large Supplementary Material:} \\
    Sparse Function-space Representation \\ of Neural Networks}
\pagestyle{empty}
